An Airdrop That Preserves Recipient Privacy

### 2019-01-18 paper/paper.tex
edits

We design a ZKPK for a $\phi$-preimage 
using a zero-knowledge protocol from~\cite[\S 3.5]{ourpaper}.
Here the verifier is given $T \in \GG^4 \times \ZZ$ 
and the prover is given $T$ and $\vv \in \ZZ^8$
where $\phi(\vv) = T$. 
The protocol works as follows:
\begin{enumerate}
\item The prover chooses
 $r_w, r_\mathit{w2}, r_\mathit{na}, r_\mathit{sa} \rgets \range{{2^{2\lambda}}}$
 and
 $r_\mathit{s1}, r_\mathit{s1w}, r_\mathit{sa}, r_\mathit{s2} \rgets \range{N \cdot 2^{2^\lambda}}$ and sets
    $\rv \deq (r_w, r_\mathit{w2}, r_\mathit{s1}, r_a, r_\mathit{na}, 
                     r_\mathit{s1w}, r_\mathit{sa}, r_\mathit{s2}) \in \ZZ^8$. \\
It computes $\Rv \gets \phi(\rv) \in \GG^4 \times \ZZ$
and sends $(c_1, c_2, \Rv)$ to the verifier.

@@

\begin{proof}
The simulator takes $(c, t)$ as input and outputs the simulated transcript
$(c_1, c_2, \Rv, \ell, \mathit{ch}, \Zv_q, \zv_\ell)$ as follows:
\begin{alg}
\> choose random $c_1, c_2 \rgets \GG,\ \ \ell \rgets \primes$, and
\ \ $\mathit{ch} \rgets \range{2^\lambda}$, \AlgSep
\lnum \label{ln:randvell}
\> choose random $\zv_\ell \rgets \range{\ell}^8$, \AlgSep
\> choose random $Z_1, Z_2, Z_3, Z_4 \rgets \GG$, \\
\lnum \label{ln:sim}
\> choose random $\hat{r}_\text{w2}, \hat{r}_\text{na} \rgets \range{2^{2\lambda}}$
and set $Z_5 \deq \lfloor (t \cdot \mathit{ch})/\ell \rfloor + \lfloor \hat{r}_\text{w2} / \ell \rfloor - \lfloor \hat{r}_\text{na} / \ell \rfloor \in \ZZ$, \\
\> set $\Zv_q \deq (Z_1, Z_2, Z_3, Z_4, Z_5)  \in \GG^4 \times \ZZ$, \AlgSep
\lnum \label{ln:computeR}
\> compute $\Rv \deq \Zv_q^\ell \cdot \phi(\zv_\ell) / T^\mathit{ch} \in \GG^4 \times \ZZ$, \AlgSep
\> output $(c_1, c_2, \Rv, \ell, \mathit{ch}, \Zv_q, \zv_\ell)$.
\end{alg}
We show the simulated transcript is distributed statistically close
to an accepting transcript of a real protocol execution.
In a real transcript we know that $c_1$ and $c_2$ are uniform in $\GG$. 
Hence, observe that the simulated $c_1, c_2, \ell, \mathit{ch}$ are all distributed 
exactly as in a real transcript.  It remains to show that the same holds
for $\zv_\ell, \Zv_q$, and $\Rv$. 
In an interaction with a real prover the quantities
$r_\mathit{s1}, r_\mathit{s1w}, r_\mathit{sa}, r_\mathit{s2}$
are chosen uniformly in $\range{N \cdot 2^{2^\lambda}}$.
Consider one such value $r \rgets \range{N \cdot 2^{2^\lambda}}$.
Because $N \cdot 2^{2\lambda}$ is much larger than $\abs{\GG} \cdot \ell$
we know that the pair
$\big(\lfloor r/\ell \rfloor \bmod \abs{\GG},\ (r \bmod \ell) \big)$
is statistically close to uniform in $\range{\,\abs{\GG}\,} \times \range{\ell}$.
Similarly,  in a real transcript
each of $r_w, r_\mathit{w2}, r_\mathit{na}, r_\mathit{sa}$
is chosen uniformly in $\range{2^{2 \lambda}}$ and therefore
each of these four values modulo $\ell$
is statistically close to uniform in $\range{\ell}$.
It follows that $\zv_\ell$, which is masked by $(\rv \bmod \ell)$
in a real transcript, is statistically close to uniform 
in $\range{\ell}^8$, as in line~\eqref{ln:randvell} in the simulation.

Next we argue that $\Zv_q \in \GG^4 \times \ZZ$ is simulated correctly.
Observe that each of the first four components of $\Zv_q$ is a group
element in $\GG$ that is randomized by
$h^{\lfloor r_\mathit{s1} / \ell \rfloor},\
 h^{\lfloor r_\mathit{s2} / \ell \rfloor},\
 h^{\lfloor r_\mathit{s1w} / \ell \rfloor},\
 h^{\lfloor r_\mathit{sa} / \ell \rfloor}$ respectively.
Because these four exponents are close to uniform in $\range{\,\abs{\GG}\,}$ and are
independent of everything else in the transcript, it follows that
in a real transcript $Z_1, Z_2, Z_3, Z_4$ are independent and close to uniform elements in $\GG$.  This is also how these elements are chosen in the simulation.

It remains to argue that $Z_5 \in \ZZ$ is simulated correctly. 
In a real transcript this $Z_5$ is computed as 
$Z_5 \deq \lfloor (\mathit{w2} \cdot \mathit{ch} + r_\text{w2})/\ell \rfloor -
          \lfloor (\mathit{na} \cdot \mathit{ch} + r_\text{na})/\ell \rfloor$.
Because $\mathit{w2} - \mathit{na} = t$, we have
that $Z_5$ in a real transcript satisfies
\[    Z_5 = \lfloor (t \cdot \mathit{ch})/\ell \rfloor + \lfloor r_\text{w2} / \ell \rfloor - \lfloor r_\text{na} / \ell \rfloor + \Delta \]
where $\Delta \in \{0, \pm 1\}$. 
As argued above, in a real transcript $\lfloor r_\text{w2} / \ell \rfloor$ 
and $\lfloor r_\text{na} / \ell \rfloor$ 
are statistically close to fresh uniform values in $\range{2^{2\lambda}/\ell}$, 
as in the simulated $Z_5$ in~\eqref{ln:sim}.
The only difference is that the real distribution is shifted by $\Delta$
where as the simulated distribution is not.  
However this $\Delta$ only negligibly affects the statistical distance
between the real and simulated distributions. 
The reason is that in both distributions, 
every possible value in the range of $Z_5$, 
except possibly at the edges, 
is obtained from the same number of values 
for the pair $(r_\text{w2}, r_\text{na})$. 
%%  would be good to prove this as a general lemma.
We thus conclude that the simulated $Z_5$ in~\eqref{ln:sim} 
is statistically close to its real distribution.

>>>>

We design a ZKPK for a $\phi$-preimage 
using a zero-knowledge protocol from~\cite[\S 3.5]{ourpaper}.
Here the verifier is given $T \in \GG^4 \times \ZZ$ 
and the prover is given $T$ and $\vv \in \ZZ^8$
where $\phi(\vv) = T$. 
The protocol works as follows:
\begin{enumerate}
\item The prover sets
    $\rv \deq (r_w, r_\mathit{w2}, r_\mathit{s1}, r_a, r_\mathit{na}, 
                     r_\mathit{s1w}, r_\mathit{sa}, r_\mathit{s2}) \in \ZZ^8$ where
\[ r_w, r_\mathit{w2}, r_\mathit{na}, r_\mathit{sa} \rgets \range{{2^{2\lambda}}} 
    \quad\text{and}\quad
   r_\mathit{s1}, r_\mathit{s1w}, r_\mathit{sa}, r_\mathit{s2} \rgets \range{N \cdot 2^{2\lambda}}.
\]
The prover computes $\Rv \gets \phi(\rv) \in \GG^4 \times \ZZ$
and sends $(c_1, c_2, \Rv)$ to the verifier.

@@

\begin{proof}
The simulator takes $(c, t)$ as input and outputs the simulated transcript
$(c_1, c_2, \Rv, \ell, \mathit{ch}, \Zv_q, \zv_\ell)$ as follows:
\begin{alg}
\> choose random $c_1, c_2 \rgets \GG,\ \ \ell \rgets \primes$, and
\ \ $\mathit{ch} \rgets \range{2^\lambda}$, \AlgSep
\lnum \label{ln:randvell}
\> choose random $\zv_\ell \rgets \range{\ell}^8$, \AlgSep
\lnum \label{ln:randZ}
\> choose random $Z_1, Z_2, Z_3, Z_4 \rgets \GG$, \\
\lnum \label{ln:sim}
\> choose random $\hat{r}_\text{w2}, \hat{r}_\text{na} \rgets \range{2^{2\lambda}}$
and set $Z_5 \deq \lfloor (t \cdot \mathit{ch})/\ell \rfloor + \lfloor \hat{r}_\text{w2} / \ell \rfloor - \lfloor \hat{r}_\text{na} / \ell \rfloor \in \ZZ$, \\
\> set $\Zv_q \deq (Z_1, Z_2, Z_3, Z_4, Z_5)  \in \GG^4 \times \ZZ$, \AlgSep
\lnum \label{ln:computeR}
\> compute $\Rv \deq \Zv_q^\ell \cdot \phi(\zv_\ell) / T^\mathit{ch} \in \GG^4 \times \ZZ$, \AlgSep
\> output $(c_1, c_2, \Rv, \ell, \mathit{ch}, \Zv_q, \zv_\ell)$.
\end{alg}
We show the simulated transcript is distributed statistically close
to an accepting transcript of a real protocol execution.
In a real transcript we know that $c_1$ and $c_2$ are uniform in $\GG$. 
Hence, observe that the simulated $c_1, c_2, \ell, \mathit{ch}$ are all distributed 
exactly as in a real transcript.  It remains to show that the same holds
for $\zv_\ell, \Zv_q$, and $\Rv$. 
In an interaction with a real prover the quantities
$r_\mathit{s1}, r_\mathit{s1w}, r_\mathit{sa}, r_\mathit{s2}$
are chosen uniformly in $\range{N \cdot 2^{2\lambda}}$.
Consider one such value $r \rgets \range{N \cdot 2^{2\lambda}}$.
Because $N \cdot 2^{2\lambda}$ is much larger than $\abs{\GG} \cdot \ell$
we know that the pair
$\big(\lfloor r/\ell \rfloor \bmod \abs{\GG},\ (r \bmod \ell) \big)$
is statistically close to uniform in $\range{\,\abs{\GG}\,} \times \range{\ell}$.
Similarly,  in a real transcript
each of $r_w, r_\mathit{w2}, r_\mathit{na}, r_\mathit{sa}$
is chosen uniformly in $\range{2^{2 \lambda}}$ and therefore
each of these four values modulo $\ell$
is statistically close to uniform in $\range{\ell}$.
It follows that $\zv_\ell$, which is masked by $(\rv \bmod \ell)$
in a real transcript, is statistically close to uniform 
in $\range{\ell}^8$, as in line~\eqref{ln:randvell} in the simulation.

Next we argue that $\Zv_q \in \GG^4 \times \ZZ$ is simulated correctly.
Observe that each of the first four components of $\Zv_q$ is a group
element in $\GG$ that is randomized by
$h^{\lfloor r_\mathit{s1} / \ell \rfloor},\
 h^{\lfloor r_\mathit{s2} / \ell \rfloor},\
 h^{\lfloor r_\mathit{s1w} / \ell \rfloor},\
 h^{\lfloor r_\mathit{sa} / \ell \rfloor}$ respectively.
Because these four exponents are close to uniform in $\range{\,\abs{\GG}\,}$ and are
independent of everything else in the transcript, it follows that
in a real transcript $Z_1, Z_2, Z_3, Z_4$ are independent and close to uniform elements in $\GG$.  This is also how these elements are chosen 
on line~\eqref{ln:randZ} in the simulation.

It remains to argue that $Z_5 \in \ZZ$ is simulated correctly. 
In a real transcript this $Z_5$ is computed as 
$Z_5 \deq \lfloor (\mathit{w2} \cdot \mathit{ch} + r_\text{w2})/\ell \rfloor -
          \lfloor (\mathit{na} \cdot \mathit{ch} + r_\text{na})/\ell \rfloor$.
Because $\mathit{w2} - \mathit{na} = t$, we have
that $Z_5$ in a real transcript satisfies
\[    Z_5 = \lfloor (t \cdot \mathit{ch})/\ell \rfloor + \lfloor r_\text{w2} / \ell \rfloor - \lfloor r_\text{na} / \ell \rfloor + \Delta \]
where $\Delta \in \{0, \pm 1\}$. 
In a real transcript $\lfloor r_\text{w2} / \ell \rfloor$ 
and $\lfloor r_\text{na} / \ell \rfloor$ 
are statistically close to fresh uniform values in $\range{2^{2\lambda}/\ell}$, 
as in the simulated $Z_5$ on line~\eqref{ln:sim}.
The only difference is that the real distribution is shifted by $\Delta$
where as the simulated distribution is not.  
However this $\Delta$ only negligibly affects the statistical distance
between the real and simulated distributions. 
The reason is that in both distributions, 
every possible value in the range of $Z_5$, 
except possibly at the edges, 
is obtained from the same number of values 
for the pair $(r_\text{w2}, r_\text{na})$. 
%%  would be good to prove this as a general lemma.
We thus conclude that the simulated $Z_5$ in~\eqref{ln:sim} 
is statistically close to its real distribution.
